### 2021-10-27 code/analysis_code/exploratory.Rmd
more exploratory

---
title: "exploratory"
author: "Ehsan"
date: "10/27/2021"
output: html_document
---

```{r}
library(ggplot2)
library(tidyverse)
library("readxl")
library(here)
```

#Read data
```{r}
data_location <- here::here("data","processed_data","processeddata.rds")
data1 <- readRDS(data_location)
```

#I want to see how many NA values are there for each column. So, I wrote the following code. For manuscript we dont need it.
```{r}
cbind(
      lapply(
             lapply(data1, is.na)
             , sum)
  )
```
#
#I want to see which columns have NA values
```{r}
df <- as.data.frame(
      cbind(
             lapply(
                   lapply(data1, is.na), sum)
        )
   )
rownames(subset(df, df$V1 != 0))
```
#the mean age and SD of mean age:
```{r}
mean(data1$Age)
sd(data1$Age)
```
#For each age group, whats their percentage?
```{r}
data1 %>% 
     group_by(Age) %>%
     summarise( percent = 100 * n() / nrow( data1) )

```
#people over 40
```{r}
data1 %>% 
     dplyr::filter(Age > 40 ) %>%
     group_by(Age) %>%
     summarise( percent = 100 * n() / nrow( data1) )

```

#percentage of male and female
```{r}
data1 %>% 
     group_by( Sex ) %>% 
     summarise( percent = 100 * n() / nrow( data1) )
```
#number of people for each educational category
```{r}
table(data1$`Highest educational level`)
```
```{r}
sum(data1$'Highest educational level' == 'Primary')
sum(data1$'Highest educational level' == 'Secondary')
sum(data1$'Highest educational level' == 'University Postgraduate degree holder')
sum(data1$'Highest educational level' == 'University Undergraduate degree holder')
```
#percentage of people for each student category
```{r}
data1 %>% 
     group_by( `Highest educational level` ) %>% 
     summarise( percent = 100 * n() / nrow( data1) )
```
#student/worker in the heathcare sector?
```{r}
table(data1$`Are you a student or worker in the healthcare sector?`)
```
#percentage of student/health worker

```{r}
data1 %>% 
     group_by( `Are you a student or worker in the healthcare sector?` ) %>% 
     summarise( percent = 100 * n() / nrow( data1) )
```

#household members?
```{r}
table(data1$`Who do you currently live with in the same household? Please select the number of persons in each age category.: 60 years and above`)
table(data1$`Who do you currently live with in the same household? Please select the number of persons in each age category.: 18 to 59 years`)
table(data1$`Who do you currently live with in the same household? Please select the number of persons in each age category.: 12 to 17 years`)
table(data1$`Who do you currently live with in the same household? Please select the number of persons in each age category.: Less than 12 years`)
```
```{r}
table(data1$`Which of the following categories best describes your current socio-economic situation?`)
```
```{r}
sum(data1$'Which of the following categories best describes your current socio-economic situation?' == 'High income category')
sum(data1$'Which of the following categories best describes your current socio-economic situation?' == 'Low income category')
sum(data1$'Which of the following categories best describes your current socio-economic situation?' == 'Lower middle income category')
sum(data1$'Which of the following categories best describes your current socio-economic situation?' == 'Upper middle income category ')
```

#percentage of income category?
```{r}
data1 %>% 
     group_by( `Which of the following categories best describes your current socio-economic situation?` ) %>% 
     summarise( percent = 100 * n() / nrow( data1) )
```
```{r}
table(data1$`Do you live in:`)
```
```{r}
sum(data1$'Do you live in:' == 'A rural place / village')
sum(data1$'Do you live in:' == 'A Sub-urban setting / urban slum')
sum(data1$'Do you live in:' == 'An urban setting / city / town')
```
#percentage of living setup?
```{r}
data1 %>% 
     group_by( `Do you live in:` ) %>% 
     summarise( percent = 100 * n() / nrow( data1) )
```
```{r}
table(data1$`Which source of COVID-19 information/advice do you trust the most?`)
```
#most trusted source?
```{r}
data1 %>% 
     group_by( `Which source of COVID-19 information/advice do you trust the most?` ) %>% 
     summarise( percent = 100 * n() / nrow( data1) )
```

```{r}
sum(data1$'Are you currently working/studying from home because of the COVID-19 outbreak?' == 'No')
sum(data1$'Are you currently working/studying from home because of the COVID-19 outbreak?' == 'Yes')
```

#percentage of work from home or not?
```{r}
data1 %>% 
     group_by( `Are you currently working/studying from home because of the COVID-19 outbreak?` ) %>% 
     summarise( percent = 100 * n() / nrow( data1) )
```
#Descriptive analysis for each question?
```{r}
table(data1$`Have you been quarantined (either at home or else where) at any point in time during the Covid-19 epidemic?`)
```
```{r}
table(data1$`Since the beginning of the COVID-19 outbreak, do you have information on your infection status?`)
```
#correct knowledge of their status
```{r}
data1 %>% 
     group_by( `Since the beginning of the COVID-19 outbreak, do you have information on your infection status?` ) %>% 
     summarise( percent = 100 * n() / nrow( data1) )
```

```{r}
table(data1$`How worried/fearful are you about becoming (re)infected by the coronavirus?: On a scale of 1 to 5`)
```
#worried about re-infection
```{r}
data1 %>% 
     group_by( `How worried/fearful are you about becoming (re)infected by the coronavirus?: On a scale of 1 to 5` ) %>% 
     summarise( percent = 100 * n() / nrow( data1) )
```

```{r}
table(data1$`Do you have any of the following chronic/underlying diseases? (many answers possible): Heart disease`)
```
```{r}
table(data1$`Do you have any of the following chronic/underlying diseases? (many answers possible): Hypertension`)
```
```{r}
table(data1$`Do you have any of the following chronic/underlying diseases? (many answers possible): Diabetes`)
```
```{r}
table(data1$`Do you have any of the following chronic/underlying diseases? (many answers possible): Cancer`)
```
```{r}
table(data1$`Do you have any of the following chronic/underlying diseases? (many answers possible): HIV`)
```
```{r}
table(data1$`Do you have any of the following chronic/underlying diseases? (many answers possible): Tuberculosis`)
```
```{r}
table(data1$`Do you have any of the following chronic/underlying diseases? (many answers possible): Asthma`)
```
```{r}
table(data1$`Do you have any of the following chronic/underlying diseases? (many answers possible): None of the above`)
```
```{r}
data1 %>% 
     group_by( `Do you have any of the following chronic/underlying diseases? (many answers possible): None of the above` ) %>% 
     summarise( percent = 100 * n() / nrow( data1) )
```

```{r}
table(data1$`During the past 7 days, have you been observing any of the following preventive measures against COVID-19? (Tick all answers that apply): Social distancing of at least 1.5m`)
```
```{r}
table(data1$`During the past 7 days, have you been observing any of the following preventive measures against COVID-19? (Tick all answers that apply): Wearing face mask`)
```
```{r}
table(data1$`During the past 7 days, have you been observing any of the following preventive measures against COVID-19? (Tick all answers that apply): Hand hygiene (regular handwashing with soap or using hand gel)`)
```
```{r}
table(data1$`During the past 7 days, have you been observing any of the following preventive measures against COVID-19? (Tick all answers that apply): Coughing hygiene (covering mouth when coughing or sneezing)`)
```
```{r}
table(data1$`During the past 7 days, have you been observing any of the following preventive measures against COVID-19? (Tick all answers that apply): None of the above`)
```
```{r}
data1 %>% 
     group_by( `During the past 7 days, have you been observing any of the following preventive measures against COVID-19? (Tick all answers that apply): None of the above` ) %>% 
     summarise( percent = 100 * n() / nrow( data1) )
```

```{r}
table(data1$`During the past two weeks, how often have you been bothered by each of the following symptoms?: Feeling down, depressed, or hopeless (low spirits)`)
```
```{r}
table(data1$`During the past two weeks, how often have you been bothered by each of the following symptoms?: Little interest or pleasure in doing things`)
```
```{r}
table(data1$`During the past two weeks, how often have you been bothered by each of the following symptoms?: Not being able to stop or control worrying`)
```
```{r}
table(data1$`During the past two weeks, how often have you been bothered by each of the following symptoms?: Feeling nervous, anxious or on edge`)
```
```{r}
table(data1$`In your opinion, can someone be re-infected with coronavirus after recovering from a previous COVID-19 infection?`)
```
```{r}
table(data1$`In your opinion, can COVID-19 infection be prevented by a vaccine?`)
```
```{r}
table(data1$`To the best of your knowledge, is there currently an effective vaccine against COVID-19?`)
```
```{r}
table(data1$`Would you be willing to take the COVID-19 vaccine when it becomes available?: If it is at least 50% effective`)
```
```{r}
table(data1$`Would you be willing to take the COVID-19 vaccine when it becomes available?: If it is at least 75% effective`)
```
```{r}
table(data1$`Would you be willing to take the COVID-19 vaccine when it becomes available?: If it is at least 90% effective`)
```
```{r}
table(data1$`Would you be willing to allow your children take the COVID-19 vaccine when it becomes available?: If it is at least 95% effective`)
```
```{r}
table(data1$`What are some of the possible reasons why you would hesitate to take the COVID-19 vaccine? (many answers possible):  I don't think COVID-19 existsists`)
```
```{r}
table(data1$`What are some of the possible reasons why you would hesitate to take the COVID-19 vaccine? (many answers possible): I think the vaccine is not effective`)
```
```{r}
table(data1$`What are some of the possible reasons why you would hesitate to take the COVID-19 vaccine? (many answers possible): I think the vaccine is designed to harm us`)
```
```{r}
table(data1$`What are some of the possible reasons why you would hesitate to take the COVID-19 vaccine? (many answers possible): I am scared of side-effects of the vaccine`)
```
```{r}
table(data1$`What are some of the possible reasons why you would hesitate to take the COVID-19 vaccine? (many answers possible): My body is naturally strong, I don't need a vaccine to fight COVID-19`)
```
```{r}
table(data1$`What are some of the possible reasons why you would hesitate to take the COVID-19 vaccine? (many answers possible): I already had COVID-19, so I think I am immune to the disease`)
```
```{r}
table(data1$`What are some of the possible reasons why you would hesitate to take the COVID-19 vaccine? (many answers possible): The COVID-19 pandemic is finished in my country, no need for a vaccine now`)
```
```{r}
table(data1$`What are some of the possible reasons why you would hesitate to take the COVID-19 vaccine? (many answers possible): None of the above`)
```
```{r}
table(data1$`What are some of the possible reasons why you would hesitate to take the COVID-19 vaccine? (many answers possible): Other reasons (please specify)`)
```
```{r}
table(data1$`Please specify other reasons for hesitating to receive the COVID-19 vaccine:`)
```
```{r}
table(data1$`How important is it to you that by getting the COVID-19 vaccine, you would protect your own health?: On a scale of 1 to 5`)
```
```{r}
table(data1$`How important is it to you that by getting the COVID-19 vaccine, you would protect the health of other people in your community?: On a scale of 1 to 5`)
```
```{r}
table(data1$`I fully understand what this study is about, and I freely consent to participate. All the information I provide can be used by researchers to better understand coronavirus disease in my country.`)
```