### 2021-10-28 code/analysis_code/analysisscript.Rmd
.

---
title: "analysis"
author: "Ehsan"
date: "10/27/2021"
output: html_document
---

```{r}
library(tidymodels)
library(readr)       # for importing data
library(tidyverse)
library(dotwhisker)
library(broom.mixed)
library(here)
```

```{r}
data_location <- here::here("data","processed_data","processeddata2.rds")
```

#load data. 
```{r}
mydata <- readRDS(data_location)
mydata
```

# this is the linear model with ordinary least square method
```{r}
is.factor(mydata$infection_status_info)
```