### 2021-10-28 products/manuscript/Manuscript.Rmd

---
title: "Using Andersen’s model of health care utilization to assess factors associated with COVID-19 testing among adults in Bangladesh: an online survey"
subtitle: ""
author: Ehsan Suez
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: article
output:
  bookdown::word_document2: 
    toc: false
    number_sections: true
  bookdown::html_document2: 
    toc: false
bibliography: ../references.bib
csl: ../apa.csl
---

The structure below is a possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed.

# Summary/Abstract
_Write a summary of your project._

# Introduction

## General Background Information
This study aimed to investigate, using the Andersen’s model of health care utilization, factors associated with COVID-19 testing among adults in Bangladesh.

## Description of data and data source
This will be a descriptive cross-sectional online study. To get data for this study, between 10 December 2020 and 9 February 2021 an online survey was organized in Bangladesh. In total 759 adults (median age 45 years, interquartile range 33-57 years, range 18-93 years), participated in the study. COVID-19 testing/infection status was assessed by self-report.

## Questions/Hypotheses to be addressed
The main question is to find out the key factors associated with covid-19 testing

# Methods and Results

_Study design, sample, and procedure_

This will be a descriptive cross-sectional online study conducted in Bangladesh between 10 December 2020 to 9 February 2021. Participant inclusion criteria were 18 years and older, any gender, and provision of electronic informed consent.

_Measures_
Using Andersen’s model of health care utilization [Andersen 1995], study variables will be categorized into outcome variable, predisposing factors, enabling/disabling factors and need for care factors.

_Outcome variable_
COVID-19 testing/infection status will be assessed with the question, “Since the beginning of the COVID-19 outbreak, do you have information on your infection status?” Response options were 1=not tested/does not know test results, 2=negative, and 3=positive.

_Predisposing factors_
Sociodemographic factors included age, sex, country of residence, educational level, and the (estimated) age(s) of their housemate(s). Chronic/underlying diseases includs heart disease, hypertension, diabetes, cancer, HIV, tuberculosis, and chronic asthma; coded as “0” none and “1” at least presence of one clinically diagnosed condition.

_COVID-19 preventive measures_
Participants were asked, “During the past 7 days, have you been observing any of the following preventive measures against COVID-19? 1) Social distancing of at least 1.5m, 2) Wearing a face mask, 3) Hand hygiene (regular handwashing with soap or using hand gel), and 4) Coughing hygiene (covering the mouth when coughing or sneezing). A composite non-adherence to all four COVID-19 preventive measure was calculated by coding each negative response with “1”, summing scores ranging from 0-4 (Cronbach’s alpha 0.7).

_Enabling/d_
Enabling factors include self-perceived socio-economic status, self-perceived area of residence, being a student or worker in the health care sector, source of COVID-19 information/advice most trusted (coded as other, including family and friends, radio/TV, social media, religious authorities and health personnel.

_Need for care factors_ include two questions on 1) the level of fear/worry of being infected with COVID-19 (ranging from 1=not at all worried to 5=extremely worried), and 2) having been quarantined (either at home or elsewhere) at any point in time during the COVID-19 epidemic.

_Data analysis_

Descriptive statistics will be used to describe the study population. Logistic regression will be used to assess associations between predisposing factors, enabling and disabling factors, need of care factors and COVID-19 testing status, COVID-19 positive versus negative status and COVID-19 positive versus negative and not tested status. Variables significant at <0.05 in univariate analyses were subsequently included in the multivariable logistic regression models. Statistical analyses will conducted using R.

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

## Data import and cleaning_ html_document
---
```{r}
library(here)
```
## Exploratory analysis

_Use a combination of text/tables/figures to explore and describe your data. You should produce plots or tables or other summary quantities for the most interesting/important quantities in your data. Depending on the total number of variables in your dataset, explore all or some of the others. FIgures produced here might be histograms or density plots, correlation plots, etc. Tables might summarize your data._

```{r summarytable,  echo=FALSE}
resulttable=readRDS("../../results/summarytable.rds")
knitr::kable(resulttable, caption = 'Age Percentage of Participants.')
```
**78.9% of total participants are less than or equal  to 35 years old.**

```{r summarytable2,  echo=FALSE}
resulttable=readRDS("../../results/summarytable2.rds")
knitr::kable(resulttable, caption = 'Sex Percentage of Participants.')
```
**While 45.18% of the participants were female, 54.82% were male.**

```{r summarytable3,  echo=FALSE}
resulttable=readRDS("../../results/summarytable3.rds")
knitr::kable(resulttable, caption = 'Education Background Percentage of Participants.')
```
**78.05% of the participants had a bachelor or post-bachelor degree**
```{r summarytable4,  echo=FALSE}
resulttable=readRDS("../../results/summarytable4.rds")
knitr::kable(resulttable, caption = 'Income Percentage of Participants.')
```
**Almost 50% of the participants belongs to "upper middle income category"**
```{r summarytable5,  echo=FALSE}
resulttable=readRDS("../../results/summarytable5.rds")
knitr::kable(resulttable, caption = 'Residence Percentage of Participants.')
```
**Less than 10% of the participants live in a village/rural area**
```{r summarytable6,  echo=FALSE}
resulttable=readRDS("../../results/summarytable6.rds")
knitr::kable(resulttable, caption = 'Source of information Percentage of Participants.')
```
**Health professionals and Tv/radio seems to be the most trusted source to the participants.**
```{r summarytable7,  echo=FALSE}
resulttable=readRDS("../../results/summarytable7.rds")
knitr::kable(resulttable, caption = 'Percentage of Participants working from home or not.')
```
**Apparently equal half of the participants worked from or not**
```{r summarytable8,  echo=FALSE}
resulttable=readRDS("../../results/summarytable8.rds")
knitr::kable(resulttable, caption = 'Percentage of Participants quarantined or not at some stage.')
```
**Major half of the participants did quarantine at some stage during the study**
```{r summarytable9,  echo=FALSE}
resulttable=readRDS("../../results/summarytable9.rds")
knitr::kable(resulttable, caption = 'Percentage of covid test status of the participants.')
```
**63% of the participants didn't go through a COVID test**
```{r summarytable10,  echo=FALSE}
resulttable=readRDS("../../results/summarytable10.rds")
knitr::kable(resulttable, caption = "Percentage of participants' worry level about (re)infection with COVID-19.")
```
**Almost half of the participants (46%) were moderate or more worried about reinfection**
```{r summarytable11,  echo=FALSE}
resulttable=readRDS("../../results/summarytable11.rds")
knitr::kable(resulttable, caption = "Percentage of participants' history of chronic disease")
```
**Only 22% of the participants recorded to have some sort of chronic disease**
_Continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in. If your dataset is small, you can do that for all variables. Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones._

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. t-tests, simple regression model with 1 predictor, etc.) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@Leek2015a] discusses types of analyses.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.

# References