### 2021-11-03 products/manuscript/supplementary.Rmd

---
title: "supplementary"
author: "Ehsan"
date: "11/3/2021"
output: html_document
---

```{r summarytable,  echo=FALSE}
resulttable=readRDS("../../results/summarytable.rds")
knitr::kable(resulttable, caption = 'Age Percentage of Participants.')
```
**78.9% of total participants are less than or equal  to 35 years old.**

```{r summarytable2,  echo=FALSE}
resulttable=readRDS("../../results/summarytable2.rds")
knitr::kable(resulttable, caption = 'Sex Percentage of Participants.')
```
**While 45.18% of the participants were female, 54.82% were male.**

```{r summarytable3,  echo=FALSE}
resulttable=readRDS("../../results/summarytable3.rds")
knitr::kable(resulttable, caption = 'Education Background Percentage of Participants.')
```
**78.05% of the participants had a bachelor or post-bachelor degree**
```{r summarytable4,  echo=FALSE}
resulttable=readRDS("../../results/summarytable4.rds")
knitr::kable(resulttable, caption = 'Income Percentage of Participants.')
```
**Almost 50% of the participants belongs to "upper middle income category"**
```{r summarytable5,  echo=FALSE}
resulttable=readRDS("../../results/summarytable5.rds")
knitr::kable(resulttable, caption = 'Residence Percentage of Participants.')
```
**Less than 10% of the participants live in a village/rural area**
```{r summarytable6,  echo=FALSE}
resulttable=readRDS("../../results/summarytable6.rds")
knitr::kable(resulttable, caption = 'Source of information Percentage of Participants.')
```
**Health professionals and Tv/radio seems to be the most trusted source to the participants.**
```{r summarytable7,  echo=FALSE}
resulttable=readRDS("../../results/summarytable7.rds")
knitr::kable(resulttable, caption = 'Percentage of Participants working from home or not.')
```
**Apparently equal half of the participants worked from or not**
```{r summarytable8,  echo=FALSE}
resulttable=readRDS("../../results/summarytable8.rds")
knitr::kable(resulttable, caption = 'Percentage of Participants quarantined or not at some stage.')
```
**Major half of the participants did quarantine at some stage during the study**
```{r summarytable9,  echo=FALSE}
resulttable=readRDS("../../results/summarytable9.rds")
knitr::kable(resulttable, caption = 'Percentage of covid test status of the participants.')
```
**63% of the participants didn't go through a COVID test**
```{r summarytable10,  echo=FALSE}
resulttable=readRDS("../../results/summarytable10.rds")
knitr::kable(resulttable, caption = "Percentage of participants' worry level about (re)infection with COVID-19.")
```
**Almost half of the participants (46%) were moderate or more worried about reinfection**
```{r summarytable11,  echo=FALSE}
resulttable=readRDS("../../results/summarytable11.rds")
knitr::kable(resulttable, caption = "Percentage of participants' history of chronic disease")
```
**Only 22% of the participants recorded to have some sort of chronic disease**