### 2021-11-12 code/analysis_code/more.Rmd

---
---
title: "analysis"
author: "Ehsan"
date: "10/27/2021"
output: html_document
---

```{r}
library(tidymodels)
library(readr)       # for importing data
library(tidyverse)
library(dotwhisker)
library(broom)
library(table1)
library(here)
library(nnet)
library(patchwork)
library(knitr)
library(GGally)
library(pacman)
library(dplyr)
```

```{r}
data_location <- here::here("data","processed_data","processeddata.rds")
```

#load data. 
```{r}
mydata <- readRDS(data_location)
View(mydata)
```

#outcome split
```{r}
mydata$infection_status_info %>% table()
```

#make a recipe

```{r}
mydata_recipe <- 
  recipe(infection_status_info ~ Age + Sex + education + economic_status + student_or_heathworker + residence + trusted_source + work_from_home + quarantined_or_not + infection_status_info + worry_level + chronic_disease_history + social_distance_practice + follow_preventive_measure, data = mydata)
```

#make a model
#choose multinom_reg() + nnet
```{r}
mydata_model <- 
  multinom_reg() %>% 
  set_engine("nnet")
```

#resampling method
```{r}
set.seed(200)
mydata_folds <- vfold_cv(mydata, v = 10)
```

#make workflow
```{r}
mydata_wf <- 
  workflow() %>% 
  add_model(mydata_model) %>% 
  add_recipe(mydata_recipe)
```

#fit
```{r}
mydata_fit <- 
 mydata_wf %>% 
  fit_resamples(mydata_folds)
```

#metrics
```{r}
collect_metrics(mydata_fit)
```