### 2021-11-12 code/analysis_code/analysisscript.Rmd
Referencing Completed!

---
title: "analysis"
author: "Ehsan"
date: "10/27/2021"
output: html_document
---

```{r}
library(tidymodels)
library(readr)       # for importing data
library(tidyverse)
library(dotwhisker)
library(broom.mixed)
library(table1)
library(here)
```

```{r}
data_location <- here::here("data","processed_data","processeddata.rds")
```

#load data. 
```{r}
mydata <- readRDS(data_location)
```

#outcome split
```{r}
mydata$infection_status_info %>% table()
```

#get a table
```{r}
overall_summary <- table1(~ Age + Sex + education + economic_status + student_or_heathworker + residence + trusted_source + work_from_home + quarantined_or_not + infection_status_info + worry_level + chronic_disease_history + social_distance_practice + follow_preventive_measure   , data=mydata, overall="Total")
overall_summary
```

```{r}
df100 = data.frame(overall_summary)
summarytable_file = here("results", "Overall.rds")
saveRDS(df100, file = summarytable_file)
```

#The main "outcome variable" is the knowledge of infection status (positive/negative/not tested). There are "predisposing factors" (age, sex, education level, chronic disease history etc), "enabling factors" (socio-economic status, area of residence, source of covid news etc) and "need for care factors" (level of fear of reinfection, quarantined or not). We'll do logistic regression analysis for key factors of epredisposing, enabling and need for care factors.

#We'll start with logistic regression model betweem infection status and Age, Sex, Education level (Key Predisposing Factors)
```{r}
lg_mod <- 
       logistic_reg() %>% 
       set_engine("glm")
lg_fit <-
       lg_mod %>%
       fit(infection_status_info ~ Age + Sex + education, data = mydata)
lg_fit
```

```{r}
tidy(lg_fit)
```

```{r}
lg_fit_table_predisposing <- tidy(lg_fit)
```

# saved it as a file
```{r}
summarytable_file12 = here("results", "lg_fit_table_predisposing.rds")
saveRDS(lg_fit_table_predisposing, file = summarytable_file12)
```

```{r}
glance(lg_fit)
```

```{r}
lg_fit_statistics_predisposing <- glance(lg_fit)
```

# saved it as a file
```{r}
summarytable_file13 = here("results", "lg_fit_statistics_predisposing.rds")
saveRDS(lg_fit_statistics_predisposing, file = summarytable_file13)
```

#Then, we'll do log model betweem infection status and socio economic status, area of residence,student/healthworker, source of covid news (Key enabling Factors)
```{r}
lg_fit1 <-
       lg_mod %>%
       fit(infection_status_info ~ economic_status + residence + trusted_source , data = mydata)
lg_fit1
```

```{r}
tidy(lg_fit1)
```

```{r}
lg_fit_table_enabling <- tidy(lg_fit1)
```

# saved it as a file
```{r}
summarytable_file14 = here("results", "lg_fit_table_enabling.rds")
saveRDS(lg_fit_table_enabling, file = summarytable_file14)
```

```{r}
glance(lg_fit1)
```

```{r}
lg_fit_statistics_enabling <- glance(lg_fit1)
```

# saved it as a file
```{r}
summarytable_file15 = here("results", "lg_fit_statistics_enabling.rds")
saveRDS(lg_fit_statistics_enabling, file = summarytable_file15)
```

#Then, we'll do log model betweem infection status and level of fear of infection and quarantined or not (Key need for care factors )
```{r}
lg_fit2 <-
       lg_mod %>%
       fit(infection_status_info ~ quarantined_or_not + worry_level , data = mydata)
lg_fit2
```

```{r}
tidy(lg_fit2)
```

```{r}
lg_fit_table_needforcare <- tidy(lg_fit2)
```

# saved it as a file
```{r}
summarytable_file16 = here("results", "lg_fit_table_needforcare.rds")
saveRDS(lg_fit_table_needforcare, file = summarytable_file16)
```

```{r}
glance(lg_fit2)
```

```{r}
lg_fit_statistics_needforcare <- glance(lg_fit2)
```

# saved it as a file
```{r}
summarytable_file17 = here("results", "lg_fit_statistics_needforcare.rds")
saveRDS(lg_fit_statistics_needforcare, file = summarytable_file17)
```

#compare three models
```{r}
compare_log_model <- anova(lg_fit$fit, lg_fit1$fit, lg_fit2$fit)
compare_log_model
```

#we want to predict each infection status level, not a single contrast. It was a bit difficult to figure out how to do this, but one needs to search here for the model to find the function one needs, turns out it is called multinom_reg()

#make a model
#choose multinom_reg() + nnet
```{r}
mydata_model <- 
  multinom_reg() %>% 
  set_engine("nnet")
```

#make a recipe

```{r}
mydata_recipe <- 
  recipe(infection_status_info ~ Age + Sex + education + economic_status + student_or_heathworker + residence + trusted_source + work_from_home + quarantined_or_not + infection_status_info + worry_level + chronic_disease_history + social_distance_practice + follow_preventive_measure, data = mydata)
```

# Put 70% of the data into the training set and 30% into testing using strata=BodyTemp
```{r}
data_split <- initial_split(mydata, prop = 0.7, strata = infection_status_info)
```

# Create data frames for the two sets:
```{r}
train_data <- training(data_split)
test_data  <- testing(data_split)
```

# training set proportions by class
```{r}
train_data %>% 
  count(infection_status_info) %>% 
  mutate(prop = n/sum(n))
```

# test set proportions by class
```{r}
test_data %>% 
  count(infection_status_info) %>% 
  mutate(prop = n/sum(n))
```
##5-fold cross validation, 5 times repeated, stratified on `infection status` for the CV folds
```{r}
mydata_folds <- vfold_cv(train_data, v = 5, repeats =5, strata= "infection_status_info")
```

#make workflow
```{r}
mydata_wf <- 
  workflow() %>% 
  add_model(mydata_model) %>% 
  add_recipe(mydata_recipe)
```

#fit
```{r}
mydata_fit <- 
 mydata_wf %>% 
  fit_resamples(mydata_folds)
```

#metrics
```{r}
collect_metrics(mydata_fit)
```